---
title: "Adding third party plots"
author: 
- name: Samuel Wieczorek
- name: Thomas Burger
package: omXplore
abstract: >
    The package omXplore is able to embed third party plots. This vignette aims to explain how to do to add an external plot.
date: "`r Sys.Date()`"
output:
    BiocStyle::html_document:
        number_sections: yes
        toc_float: true
vignette: >
    %\VignetteIndexEntry{Adding third party plots}
    %%\VignetteKeywords{Softmware}
    %\VignetteEncoding{UTF-8}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r env, message = FALSE, warning = FALSE, echo = FALSE}
library("omXplore")
```

# Introduction

One interesting feature about `omXplore` is the capacity to insert custom plots into the collection of vignettes of the Shiny app `view_dataset()`.



# Adding external plots

omXplore can plot plots coming from third packages. For that purpose, two points may be considered:

-   make a shiny app available in the global R environment

-   ensure that the name of the two functions `ui()` and `server()` of
    the app respect a predefined syntax. More specifically, the name of
    the shiny app is composed by the prefix 'mod_ds\_' (which stands for
    'descriptive statistics') and a user-defined name. As for each Shiny
    app, there must be two functions (one for the ui part and the second
    one for the server part.). Thus, in this example, the shiny app
    called `foo` must provide the functions `foo_ui()` and
    `foo_server()`.

In this case, a clickable button is added to the series of vignettes (at the top of the right side
of the UI). If an image has been correctly configured, it is shown in this button.

The use of those external apps is devoted to its developer, the
interface of `omXplore` is only a shell that presents apps in a
user-friendly interface.

To add external app to `omXplore`, please use the global function
`view_dataset()` as follows. In the example below, one want to add the
app called `extFoo1` and `extFoo2` from the package `omXplore`. It is
only for the example, this app contains a simple plot just to
demonstrate how it works

```         
#data(vdata)
#addons <- list(omXplore = c("extFoo1", "extFoo2"))
#view_dataset(vdata, addons)
```

If there are icons to display within omXplore, the png images must be
stored in the 'images' directory of the corresponding package.

## Technical point of view

# Session information

```{r}
sessionInfo()
```
